### 2022-04-21 final_manuscript_no_markup.tex

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing during the organization of useR! 2021. HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR! conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups. The R Foundation is currently funding YBS to support the useR! 2022 organization team. HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
The organization of useR! 2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles. RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively.

For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission.

>>>>

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS. HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.

The organization of useR! 2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.

\section*{Data availability}
For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any author accepted manuscript version arising from this submission. The manuscript, figures, supplementary material, and supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.
